---
title: "01_Wrangling"
author: "Matheus Ardigueri"
date: "2026-01-27"
output: html_document
---

# 1) Setup
```{r Loading Packages, message=FALSE, warning=FALSE}
#install.packages("glmnet")

library(readxl) # to read excel files
library(janitor) # to clean data
library(dplyr) # wrangling
library(tidyr) # wrangling
library(readr)
library(tidyverse)
library(ggplot2)
library(glmnet)
library(randomForest)
library(readxl)
library(dplyr)
library(tidyr)
library(purrr)
library(stringr)
library(janitor)
#install.packages("corrplot")
library(corrplot)
library(ggplot2)
library(gridExtra)
library(ggcorrplot)
library(caret)
#install.packages("ranger")
library(ranger)
library(lubridate)
```

```{r Loading Dataset with images, message=FALSE, warning=FALSE}

ponder <- read.csv("../Data/2024_reflectance.csv", sep = ";")


ponder
```

```{r Cleaning Colum Names}
clean_ponder <- ponder %>%
  clean_names() %>%
  mutate(across(8:ncol(.), as.numeric)) %>%
  select(-cultivar_id)

clean_ponder

```

# 2) Data view 
# 2.1) Boxplots

```{r Combined Plot For Dates Collected in 2024}

# Main plot
phys_vars <- c("ca", "cb", "cx_c", "pn", "gs", "etr")

ponder_long_2024 <- clean_ponder %>%
  mutate(across(all_of(phys_vars), as.numeric)) %>%
  mutate(dap = as.factor(dap)) %>%
  pivot_longer(
    cols = all_of(phys_vars),
    names_to = "variable",
    values_to = "value"
  ) %>%
  filter(year == 2024)
  

variable_labels_f <- c(
  gs = "Stomata Cond. (mol m-2s-1)",
  pn = "Net Photo (µmol m-2s-1)",
  etr = "ETR (µmol m-2s-1)",
  ca = "Chl.a (µg/cm-2)",
  cb = "Chl.b (µg/cm-2)",
  cx_c = "Car (µg/cm-2)"
)

bp_physilog_only_selected_dates <- ggplot(
  ponder_long_2024,
  aes(
    x = dap,
    y = value,
    fill = treatment,
    group = interaction(dap, treatment)
  )
) +
  geom_boxplot() +
  facet_wrap(~ variable, scales = "free_y",
             labeller = labeller(variable = variable_labels_f))+
  scale_fill_viridis_d(name = "Treatments") +
  xlab("Days After Planting") +
  labs(fill = "Treatments") +
  ylab("Physiological Variables Values")+
  theme_light() +
  theme(
  strip.text = element_text(size = 18, colour = "black", face = "bold"), 
  legend.position = c(0.945,0.59), 
  legend.background = element_blank(),
  axis.text.x = element_text(size = 17, face = "bold", margin = ggplot2::margin(b = 10), angle = 90),
  axis.text.y = element_text(size = 18, face = "bold", margin = ggplot2::margin(l = 10)),
  axis.title.x = element_text(size = 20, face = "bold", margin = ggplot2::margin(t = 15)),
  axis.title.y = element_text(size = 20, face = "bold", margin = ggplot2::margin(r = 15)),
  legend.title = element_text(size = 17, face = "bold"),
  legend.text = element_text(size = 17, face = "bold"))

    

# Combine plots


bp_physilog_only_selected_dates

#Save

# ggsave(
#  filename = "../Output/Boxplots/2024_bp_phys.png",
#   plot = bp_physilog_only_selected_dates,
#   dpi = 300,
#   height = 10,
#   width = 16
# )

```

```{r Combined Plot For Dates Collected in 2025}
ponder_long_2025 <- clean_ponder %>%
  mutate(across(all_of(phys_vars), as.numeric)) %>%
  mutate(dap = as.factor(dap)) %>%
  pivot_longer(
    cols = all_of(phys_vars),
    names_to = "variable",
    values_to = "value"
  ) %>%
  filter(year == 2025)

bp_physilog_only_selected_dates_25 <- ggplot(
  ponder_long_2025,
  aes(
    x = dap,
    y = value,
    fill = treatment,
    group = interaction(dap, treatment)
  )
) +
  geom_boxplot() +
  facet_wrap(~ variable, scales = "free_y",
             labeller = labeller(variable = variable_labels_f))+
  scale_fill_viridis_d(name = "Treatments") +
  xlab("Days After Planting") +
  labs(fill = "Treatments") +
  ylab("Physiological Variables Values")+
  theme_light() +
  theme(
  strip.text = element_text(size = 18, colour = "black", face = "bold"), 
  legend.position = c(0.945,0.59), 
  legend.background = element_blank(),
  axis.text.x = element_text(size = 17, face = "bold", margin = ggplot2::margin(b = 10), angle = 90),
  axis.text.y = element_text(size = 18, face = "bold", margin = ggplot2::margin(l = 10)),
  axis.title.x = element_text(size = 20, face = "bold", margin = ggplot2::margin(t = 15)),
  axis.title.y = element_text(size = 20, face = "bold", margin = ggplot2::margin(r = 15)),
  legend.title = element_text(size = 17, face = "bold"),
  legend.text = element_text(size = 17, face = "bold"))

    

# Combine plots


bp_physilog_only_selected_dates_25

# Save
# ggsave(
#  filename = "../Output/Boxplots/2025_bp_phys.png",
#   plot = bp_physilog_only_selected_dates_25,
#   dpi = 300,
#   height = 10,
#   width = 16
# )
```

```{r Wrangling to combine cultivars and analyze individual relations}
ponder_cult_spec <- clean_ponder %>%
  pivot_longer(
    cols = all_of(phys_vars),
    names_to = "variable",
    values_to = "value"
  )

names(ponder_long_2024$genotype)
```

```{r Boxplot phys x cultivars}

bp_phys_cult <- ggplot(
  ponder_long_2024,
  aes(
    x = genotype,
    y = value,
    fill = treatment,
    group = interaction(genotype, treatment)
  )
) +
  geom_boxplot() +
  facet_wrap(~ variable, scales = "free_y",
             labeller = labeller(variable = variable_labels_f))+
  scale_fill_viridis_d(name = "Treatments") +
  xlab("Genotype") +
  labs(fill = "Treatments") +
  ylab("Physiological Variables Values")+
  theme_light() +
  theme(
  strip.text = element_text(size = 18, colour = "black", face = "bold"), 
  legend.position = c(0.945,0.59), 
  legend.background = element_blank(),
  axis.text.x = element_text(size = 17, face = "bold", margin = ggplot2::margin(b = 10), angle = 90),
  axis.text.y = element_text(size = 18, face = "bold", margin = ggplot2::margin(l = 10)),
  axis.title.x = element_text(size = 20, face = "bold", margin = ggplot2::margin(t = 15)),
  axis.title.y = element_text(size = 20, face = "bold", margin = ggplot2::margin(r = 15)),
  legend.title = element_text(size = 17, face = "bold"),
  legend.text = element_text(size = 17, face = "bold"))



bp_phys_cult

# ggsave(bp_phys_cult,
#        filename= "../Output/Boxplots/bp_phys_cult_2024.png",
#        height = 10,
#        width = 16)



```

```{r}
bp_phys_cults25 <- ggplot(
  ponder_long_2025,
  aes(
    x = genotype,
    y = value,
    fill = treatment,          # DAP as discrete for legend/colors
    group = interaction( genotype, treatment)
  )
) +
  geom_boxplot(
    linewidth = 1,
    position = position_dodge2(width = 0.85, preserve = "single")
  ) +
  facet_wrap(
    ~ variable,
    scales = "free_y",
    labeller = labeller(variable = variable_labels_f)
  ) +
  scale_fill_viridis_d(name = "Treatment") +
  guides(
    linetype = guide_legend(order = 2)
  ) +
  xlab("Genotype") +
  ylab("Physiological Variables Values") +
  theme_light() +
  theme(
    strip.text = element_text(size = 18, colour = "black", face = "bold"),
    legend.position = c(0.15, 0.07),
    legend.direction = "vertical",
    legend.box = "vertical",
    legend.background = element_blank(),
    legend.title = element_text(size = 17, face = "bold"),
    legend.text  = element_text(size = 17, face = "bold"),
    axis.text.x  = element_text(size = 17, face = "bold",
                                margin = ggplot2::margin(b = 10), angle = 90),
    axis.text.y  = element_text(size = 18, face = "bold",
                                margin = ggplot2::margin(l = 10)),
    axis.title.x = element_text(size = 20, face = "bold",
                                margin = ggplot2::margin(t = 15)),
    axis.title.y = element_text(size = 20, face = "bold",
                                margin = ggplot2::margin(r = 15))
  )

bp_phys_cults25

# ggsave(
#   bp_phys_cults25,
#   filename = "../Output/Boxplots/bp_phys_cults255.png",
#        height = 10,
#        width = 16
# )
```

```{r}
bp_phys_cults24 <- ggplot(
  ponder_long_2024,
  aes(
    x = genotype,
    y = value,
    fill = factor(dap),          # DAP as discrete for legend/colors
    linetype = treatment,
    group = interaction(dap, genotype, treatment)
  )
) +
  geom_boxplot(
    linewidth = 1,
    position = position_dodge2()
  ) +
  facet_wrap(
    ~ variable,
    scales = "free_y",
    labeller = labeller(variable = variable_labels_f)
  ) +
  scale_fill_viridis_d(name = "DAP") +
  scale_linetype_discrete(name = "Treatment") +
  guides(
    fill     = guide_legend(order = 1),
    linetype = guide_legend(order = 2)
  ) +
  xlab("Genotype") +
  ylab("Physiological Variables Values") +
  theme_light() +
  theme(
    strip.text = element_text(size = 18, colour = "black", face = "bold"),
    legend.position = c(0.15, 0.07),
    legend.direction = "horizontal",
    legend.box = "horizontal",
    legend.background = element_blank(),
    legend.title = element_text(size = 17, face = "bold"),
    legend.text  = element_text(size = 17, face = "bold"),
    axis.text.x  = element_text(size = 17, face = "bold",
                                margin = ggplot2::margin(b = 10), angle = 90),
    axis.text.y  = element_text(size = 18, face = "bold",
                                margin = ggplot2::margin(l = 10)),
    axis.title.x = element_text(size = 20, face = "bold",
                                margin = ggplot2::margin(t = 15)),
    axis.title.y = element_text(size = 20, face = "bold",
                                margin = ggplot2::margin(r = 15))
  )

bp_phys_cults24

# ggsave(
#   bp_phys_cults24,
#   filename = "../Output/Boxplots/bp_phys_cults24.png",
#        height = 11,
#        width = 18,
#   dpi = 900
# )
```


```{r Combined Plot for All Dates}
# bp_all_physiolog_dates <- ggplot(
#   ponder_long_p,
#   aes(
#     x = dap,
#     y = value,
#     fill = drought_treatment,
#     group = interaction(dap, drought_treatment)
#   )
# ) +
#   geom_boxplot() +
#   
#   facet_wrap(~ variable, scales = "free_y", labeller = labeller(variable = variable_labels_p))+
#   scale_color_viridis_d() +
#   xlab("Days After Planting") +
#   labs(color = "Stress Situation", fill = "Treatments") +
#   ylab("Physiological Variables Values")+
#   
#   theme_light() +
#   theme(
#     strip.text = element_text(size = 18, colour = "black", face = "bold"), 
#   legend.position = c(0.94,0.09), 
#   legend.background = element_blank(),
#   axis.text.x = element_text(size = 17, face = "bold", margin = ggplot2::margin(b = 10)),
#   axis.text.y = element_text(size = 18, face = "bold", margin = ggplot2::margin(l = 10)),
#   axis.title.x = element_text(size = 20, face = "bold", margin = ggplot2::margin(t = 15)),
#   axis.title.y = element_text(size = 20, face = "bold", margin = ggplot2::margin(r = 15)),
#   legend.title = element_text(size = 17, face = "bold"),
#   legend.text = element_text(size = 17, face = "bold"))
#     
# bp_all_physiolog_dates

# ggsave(
#   filename = "../Output/Boxplots/Physiology_alldates_boxplots.png",
#   plot = bp_all_physiolog_dates,
#   dpi = 300,
#   width = 13,
#   height = 10
# )
```

# 3) Spectral Curves
```{r Converting Bands to Wavelength and Grouping by to Calculate Mean}
wavelengths <- seq(400, 1000, length.out = 300)

df_avg_24 <- clean_ponder %>%
  filter(year == 2024) %>%
  pivot_longer(
    cols = starts_with("b_"),
    names_to = "band",
    values_to = "reflectance"
  ) %>%
  mutate(
    band_num = as.numeric(str_remove(band, "b_")),
    wavelength = wavelengths[band_num]
  ) %>% 
  group_by(treatment, genotype, wavelength) %>%
  summarise(
    mean_reflectance = mean(reflectance, na.rm = TRUE),
    .groups = "drop"
  )

df_avg_25 <- clean_ponder %>%
  filter(year == 2025) %>%
  pivot_longer(
    cols = starts_with("b_"),
    names_to = "band",
    values_to = "reflectance"
  ) %>%
  mutate(
    band_num = as.numeric(str_remove(band, "b_")),
    wavelength = wavelengths[band_num]
  ) %>% 
  group_by(treatment, genotype, wavelength) %>%
  summarise(
    mean_reflectance = mean(reflectance, na.rm = TRUE),
    .groups = "drop"
  )


```

```{r Naming Different Contrasted Colors}
cultivar_colors <- c(
  "#0072B2",  # blue
  "#E69F00",  # orange
  "#009E73",  # bluish green
  "#56B4E9",  # sky blue
  "#D55E00",  # vermillion
  "#CC79A7",  # purple
  "#F0E442",  # yellow (safe contrast)
  "#999999",  # gray
  "black"  ) # black

all_cultivars <- sort(unique(c(
  df_avg_24$genotype,
  df_avg_25$genotype
)))

# If you don't have enough colors, repeat or expand palette
if (length(all_cultivars) > length(cultivar_colors)) {
  cultivar_colors <- rep(cultivar_colors, length.out = length(all_cultivars))
}

# Assign names consistently
names(cultivar_colors) <- all_cultivars
```

```{r Creating Reflectance Plot plot for each different cultivar 2024}
# B = 400
# G = 550
# R = 660

p_ref_facet_cultivar_24 <- df_avg_24 %>%
  ggplot(aes(
    x = wavelength,
    y = mean_reflectance,
    color = treatment
  )) +
  geom_line(linewidth = 0.7) +
  geom_vline(xintercept = 450, colour = "blue") +
  geom_vline(xintercept = 550, color = "green") +
  geom_vline(xintercept = 660, color = "red") +
  scale_color_viridis_d() +
  facet_wrap(~ genotype, nrow = 3, ncol = 3) +
  labs(
    title = element_blank(),
    x = "Wavelength (nm)",
    y = "Reflectance values",
    color = "Treatment"
  ) +
  theme_gray() +
  theme(
    strip.text = element_text(size = 18, colour = "black", face = "bold"), 
  legend.position = c(0.92, 0.12), 
  legend.background = element_blank(),
  ylab("Reflectance"),
  xlab("Wavelength (nm)"),
  axis.text.x = element_text(size = 20, face = "bold", margin = ggplot2::margin(b = 10)),
  axis.text.y = element_text(size = 20, face = "bold", margin = ggplot2::margin(l = 10)),
  axis.title.x = element_text(size = 22, face = "bold", margin = ggplot2::margin(t = 15)),
  axis.title.y = element_text(size = 22, face = "bold", margin = ggplot2::margin(r = 15)),
  legend.title = element_text(size = 22, face = "bold"),
  legend.text = element_text(size = 20, face = "bold"))

p_ref_facet_cultivar_24

# ggsave(
#   plot = p_ref_facet_cultivar_24,
#   filename = "../Output/Spectral/p_ref_facet_cultivar_24.png",
#   dpi = 600,
#   width = 16,
#   height = 13
# )

```

```{r Creating Reflectance Plot plot for each different cultivar 2025}
p_ref_facet_cultivar_25 <- df_avg_25 %>%
  ggplot(aes(
    x = wavelength,
    y = mean_reflectance,
    color = treatment
  )) +
  geom_line(linewidth = 0.7) +
  
  scale_color_viridis_d() +
  facet_wrap(~ genotype, nrow = 3, ncol = 3) +
  labs(
    title = element_blank(),
    x = "Wavelength (nm)",
    y = "Reflectance values",
    color = "Treatment"
  ) +
  theme_gray() +
  geom_vline(xintercept = 450, colour = "blue") +
  geom_vline(xintercept = 550, color = "green") +
  geom_vline(xintercept = 660, color = "red") +
  theme(
    strip.text = element_text(size = 18, colour = "black", face = "bold"), 
  legend.position = c(0.92, 0.12), 
  legend.background = element_blank(),
  ylab("Reflectance"),
  xlab("Wavelength (nm)"),
  axis.text.x = element_text(size = 20, face = "bold", margin = ggplot2::margin(b = 10)),
  axis.text.y = element_text(size = 20, face = "bold", margin = ggplot2::margin(l = 10)),
  axis.title.x = element_text(size = 22, face = "bold", margin = ggplot2::margin(t = 15)),
  axis.title.y = element_text(size = 22, face = "bold", margin = ggplot2::margin(r = 15)),
  legend.title = element_text(size = 22, face = "bold"),
  legend.text = element_text(size = 20, face = "bold"))

p_ref_facet_cultivar_25

# ggsave(
#   plot = p_ref_facet_cultivar_25,
#   filename = "../Output/Spectral/p_ref_facet_cultivar_25.png",
#   dpi = 600,
#   width = 16,
#   height = 13
# )
```


```{r Creating Reflectance Plot plot for each different cultivar zoomed in 24}
p_spec_zoom_24 <- df_avg_24 %>%
  ggplot(aes(
    x = wavelength,
    y = mean_reflectance,
    color = treatment
  )) +
  geom_line(linewidth = 0.7) +
  scale_color_viridis_d() +
  facet_wrap(~ genotype, nrow = 3, ncol = 3) +
  labs(
    title = element_blank(),
    x = "Wavelength (nm)",
    y = "Reflectance values",
    color = "Treatment"
  ) +
  coord_cartesian(xlim = c(740, 1000)) +
  theme_gray() +
  theme(
    strip.text = element_text(size = 18, colour = "black", face = "bold"), 
  legend.position = c(0.92, 0.12), 
  legend.background = element_blank(),
  ylab("Reflectance"),
  xlab("Wavelength (nm)"),
  axis.text.x = element_text(size = 20, face = "bold", margin = ggplot2::margin(b = 10)),
  axis.text.y = element_text(size = 20, face = "bold", margin = ggplot2::margin(l = 10)),
  axis.title.x = element_text(size = 22, face = "bold", margin = ggplot2::margin(t = 15)),
  axis.title.y = element_text(size = 22, face = "bold", margin = ggplot2::margin(r = 15)),
  legend.title = element_text(size = 22, face = "bold"),
  legend.text = element_text(size = 20, face = "bold"))

p_spec_zoom_24

# ggsave(
#   plot = p_spec_zoom_24,
#   filename = "../Output/Spectral/p_spec_zoom_24.png",
#   dpi = 600,
#   width = 16,
#   height = 13
# )

```

```{r Creating Reflectance Plot plot for each different cultivar zoomed in 25}
p_spec_zoom_25 <- df_avg_25 %>%
  ggplot(aes(
    x = wavelength,
    y = mean_reflectance,
    color = treatment
  )) +
  geom_line(linewidth = 0.7) +
  scale_color_viridis_d() +
  facet_wrap(~ genotype, nrow = 3, ncol = 3) +
  labs(
    title = element_blank(),
    x = "Wavelength (nm)",
    y = "Reflectance values",
    color = "Treatment"
  ) +
  coord_cartesian(xlim = c(740, 1000)) +
  theme_gray() +
  theme(
    strip.text = element_text(size = 18, colour = "black", face = "bold"), 
  legend.position = c(0.92, 0.12), 
  legend.background = element_blank(),
  ylab("Reflectance"),
  xlab("Wavelength (nm)"),
  axis.text.x = element_text(size = 20, face = "bold", margin = ggplot2::margin(b = 10)),
  axis.text.y = element_text(size = 20, face = "bold", margin = ggplot2::margin(l = 10)),
  axis.title.x = element_text(size = 22, face = "bold", margin = ggplot2::margin(t = 15)),
  axis.title.y = element_text(size = 22, face = "bold", margin = ggplot2::margin(r = 15)),
  legend.title = element_text(size = 22, face = "bold"),
  legend.text = element_text(size = 20, face = "bold"))

p_spec_zoom_25

# ggsave(
#   plot = p_spec_zoom_25,
#   filename = "../Output/Spectral/p_spec_zoom_25.png",
#   dpi = 600,
#   width = 16,
#   height = 13
# )

```

```{r Separating treatment in plots and cultivars as HUE}
p_combined_irrigation_2_plots <- df_avg_24 %>%
  ggplot(aes(
    x = wavelength,
    y = mean_reflectance,
    color = genotype
  )) +
  geom_line(linewidth = 0.5) +
  facet_wrap(~ treatment, nrow = 1) +
  scale_color_manual(values = cultivar_colors) +
  labs(
    title = element_blank(),
    x = "Wavelength (nm)",
    y = "Reflectance values",
    color = "Cultivar"
  ) +
  coord_cartesian(xlim = c(740, 1000)) +   # Zoom into Red Edge + NIR (Remove for whole graph)
  theme_light() +
  theme(
    strip.text = element_text(size = 18, colour = "black", face = "bold"), 
  legend.position = c(0.89, 0.20), 
  legend.background = element_blank(),
  axis.text.x = element_text(size = 17, face = "bold", margin = ggplot2::margin(b = 10)),
  axis.text.y = element_text(size = 18, face = "bold", margin = ggplot2::margin(l = 10)),
  axis.title.x = element_text(size = 20, face = "bold", margin = ggplot2::margin(t = 15)),
  axis.title.y = element_text(size = 20, face = "bold", margin = ggplot2::margin(r = 15)),
  legend.title = element_text(size = 17, face = "bold"),
  legend.text = element_text(size = 17, face = "bold"))

p_combined_irrigation_2_plots

# ggsave(
#   filename = "../Output/Spectral_plots/Reflectance_curves_by_treatments_zoom_NIR.png",
#   plot = p_combined_irrigation_2_plots,
#   dpi = 600,
#   width = 13,
#   height = 10
# )

```
```{r Suport function and setup to spectral graphs by date.}
make_mean_spectra <- function(df, target_year, wavelengths_vec) {
  df %>%
    filter(year == target_year) %>%
    pivot_longer(
      cols = starts_with("b_"),
      names_to = "band",
      values_to = "reflectance"
    ) %>%
    mutate(
      band_num = as.numeric(str_remove(band, "b_")),
      wavelength = wavelengths_vec[band_num]
    ) %>%
    group_by(year, dap, treatment, genotype, wavelength) %>%
    summarise(
      mean_reflectance = mean(reflectance, na.rm = TRUE),
      .groups = "drop"
    )
}

spectra_mean_2024 <- make_mean_spectra(clean_ponder, 2024, wavelengths)
spectra_mean_2025 <- make_mean_spectra(clean_ponder, 2025, wavelengths)

spectra_mean_2024 <- spectra_mean_2024 %>%
  filter(dap != 75) %>%
  droplevels()
spectra_mean_2024 <- spectra_mean_2024 %>%
  mutate(dap = factor(dap, levels = sort(unique(dap))))

spectra_mean_2025 <- spectra_mean_2025 %>%
  mutate(dap = factor(dap, levels = sort(unique(dap))))
```

```{r Combined Grid 24}
p_combined_irrigation_grid_2024 <- spectra_mean_2024 %>%
  ggplot(aes(
    x = wavelength,
    y = mean_reflectance,
    color = genotype
  )) +
  geom_line(linewidth = 0.5) +
  facet_grid(treatment ~ dap,
             labeller = labeller(
    dap = function(x) paste("DAP", x)
  )) +
  scale_color_manual(values = cultivar_colors) +
  coord_cartesian(xlim = c(400, 1000)) +
  theme_gray() +
  labs(
    title = element_blank(),
    x = "Wavelength (nm)",
    y = "Reflectance values",
    color = "Genotype"
  ) +
  theme(
    strip.text = element_text(size = 18, colour = "black", face = "bold"), 
  legend.position = c(0.925, 0.12), 
  legend.background = element_blank(),
  axis.text.x = element_text(size = 17, face = "bold", margin = ggplot2::margin(b = 10)),
  axis.text.y = element_text(size = 18, face = "bold", margin = ggplot2::margin(l = 10)),
  axis.title.x = element_text(size = 20, face = "bold", margin = ggplot2::margin(t = 15)),
  axis.title.y = element_text(size = 20, face = "bold", margin = ggplot2::margin(r = 15)),
  legend.title = element_text(size = 17, face = "bold"),
  legend.text = element_text(size = 17, face = "bold"))

p_combined_irrigation_grid_2024

# ggsave(plot = p_combined_irrigation_grid_2024,
#        filename = "../Output/Spectral/combined_irrigation_DAP_24.png",
#        dpi = 600,
#        width = 18,
#       height = 13
#        )

```

```{r Combined Grid 25}
p_combined_irrigation_grid_2025 <- spectra_mean_2025 %>%
  ggplot(aes(
    x = wavelength,
    y = mean_reflectance,
    color = genotype
  )) +
  geom_line(linewidth = 0.5) +
  labs(
    title = element_blank(),
    x = "Wavelength (nm)",
    y = "Reflectance values",
    color = "Genotype"
  ) +
  facet_grid(treatment ~ dap,
             labeller = labeller(
    dap = function(x) paste("DAP", x)
  )) +
  scale_color_manual(values = cultivar_colors) +
  coord_cartesian(xlim = c(400, 1000)) +
  theme_gray() +
  theme(
    strip.text = element_text(size = 18, colour = "black", face = "bold"), 
  legend.position = c(0.925, 0.14), 
  legend.background = element_blank(),
  axis.text.x = element_text(size = 17, face = "bold", margin = ggplot2::margin(b = 10)),
  axis.text.y = element_text(size = 18, face = "bold", margin = ggplot2::margin(l = 10)),
  axis.title.x = element_text(size = 20, face = "bold", margin = ggplot2::margin(t = 15)),
  axis.title.y = element_text(size = 20, face = "bold", margin = ggplot2::margin(r = 15)),
  legend.title = element_text(size = 17, face = "bold"),
  legend.text = element_text(size = 17, face = "bold"))

p_combined_irrigation_grid_2025

# ggsave(plot = p_combined_irrigation_grid_2025,
#        filename = "../Output/Spectral/combined_irrigation_DAP_25.png",
#        dpi = 600,
#        width = 18,
#       height = 13
#        )

```

## Calculating vegetation indexes to correlation.

Chlorophyll pigments:
Chlorophyll indices - (https://doi.org/10.1078/0176-1617-00887) 
CIGreen - (RNIR/R520–585)-1
REGCI - (RNIR/R695–740)-1 
MCARI - [(R7002R670)20.2(R7002R550)·(R700/R670) (https://ntrs.nasa.gov/api/citations/19950010604/downloads/19950010604.pdf) *minimizing the effects of nonphotosynthetic materials on spectral estimates of absorbed photo synthetically active radiation.*


#4) Feature Selection and Correlation.

```{r Define response variables for 24}
phys_vars <- c("gs", "pn","etr","ca","cb","cx_c")

clean_ponder_na <- clean_ponder %>%
  drop_na()
  
X <- clean_ponder_na %>%
  select(starts_with("b_")) %>%
  as.matrix()

X <- scale(X)
```

```{r Run Elastic Net (α = 0.5) for each trait}
enet_results <- map_df(phys_vars, function(var) {
  
  y <- clean_ponder_na[[var]]
  
  cv_enet <- cv.glmnet(
    x = X,
    y = y,
    alpha = 0.9,          # Elastic Net
    standardize = TRUE,
    nfolds = 20
  )
  
  coef_mat <- coef(cv_enet, s = "lambda.min")
  
  tibble(
    band = rownames(coef_mat),
    coefficient = as.numeric(coef_mat),
    variable = var
  ) %>%
    filter(band != "(Intercept)") %>%
    mutate(
      abs_coef = abs(coefficient),
      band_num = as.numeric(str_remove(band, "b_"))
    )
})

# write_csv(enet_results,'../Output/enet_results.csv')

```


```{r Select important bands}
selected_bands <- enet_results %>%
  filter(abs_coef != 0)

```

```{r Converting Wavelength}
band_to_wavelength <- function(band_num,
                               wl_min = 400,
                               wl_max = 1000,
                               n_bands = 300) {
  wl_min + (band_num - 1) * (wl_max - wl_min) / (n_bands - 1)
}
```

```{r Selecting >0 important bands and assigning regions}
selected_bands_wl <- selected_bands %>%
  mutate(
    wavelength = round(band_to_wavelength(band_num),0)
  ) %>%
  mutate(
    region = case_when(
      wavelength >= 400 & wavelength <  520 ~ "Blue",
      wavelength >= 520 & wavelength <  630 ~ "Green",
      wavelength >= 630 & wavelength <  680 ~ "Red",
      wavelength >= 680 & wavelength <  750 ~ "RedEdge",
      wavelength >= 750 & wavelength <= 1000 ~ "NIR"
      
    )
  ) 
```

```{r}
select_bcw <- selected_bands_wl %>%
  pivot_wider(names_from = band,
              values_from = coefficient)
```
```{r Getting Bands from EN}

gs_bands_chr <- paste0("b_", gs_bands)


gs_bands <- selected_bands_wl %>%
  filter(variable == "gs") %>%
  pull(band_num)

pn_bands <- selected_bands_wl %>%
  filter(variable == "pn") %>%
  pull(band_num)

pn_bands_chr <- paste0("b_", pn_bands)

etr_bands <- selected_bands_wl %>%
  filter(variable == "etr") %>%
  pull(band_num)

etr_bands_chr <- paste0("b_", etr_bands)

ca_bands <- selected_bands_wl %>%
  filter(variable == "ca") %>%
  pull(band_num)

ca_bands_chr <- paste0("b_", ca_bands)

cb_bands <- selected_bands_wl %>%
  filter(variable == "cb") %>%
  pull(band_num)

cb_bands_chr <- paste0("b_", cb_bands)

cx_c_bands <- selected_bands_wl %>%
  filter(variable == "cx_c") %>%
  pull(band_num)
cx_c_bands_chr <- paste0("b_", cx_c_bands)

```

```{r Filtering Bands directly from ponder long}
ponder_gs <- ponder_long %>%
  filter(variable == "gs",
         band %in% gs_bands_chr)%>%
  pivot_wider(names_from = variable,
              values_from = var_val) %>%
  pivot_wider(names_from = band,
              values_from = reflectance)

ponder_pn <- ponder_long %>%
  filter(variable == "pn",
         band %in% pn_bands_chr)%>%
  pivot_wider(names_from = variable,
              values_from = var_val) %>%
  pivot_wider(names_from = band,
              values_from = reflectance)

ponder_etr <- ponder_long %>%
  filter(variable == "etr",
         band %in% etr_bands_chr)%>%
  pivot_wider(names_from = variable,
              values_from = var_val) %>%
  pivot_wider(names_from = band,
              values_from = reflectance)

ponder_ca <- ponder_long %>%
  filter(variable == "ca",
         band %in% ca_bands_chr)%>%
  pivot_wider(names_from = variable,
              values_from = var_val) %>%
  pivot_wider(names_from = band,
              values_from = reflectance)

ponder_cb <- ponder_long %>%
  filter(variable == "cb",
         band %in% cb_bands_chr)%>%
  pivot_wider(names_from = variable,
              values_from = var_val) %>%
  pivot_wider(names_from = band,
              values_from = reflectance)

ponder_cx_c <- ponder_long %>%
  filter(variable == "cx_c",
         band %in% cx_c_bands_chr)%>%
  pivot_wider(names_from = variable,
              values_from = var_val) %>%
  pivot_wider(names_from = band,
              values_from = reflectance)
```

blue    <- intersect(c("b_3","b_4","b_19","b_23"), gs_bands)
green   <- intersect(c("b_70","b_76"), gs_bands)
red     <- intersect(c("b_118","b_127"), gs_bands)
rededge <- intersect(c("b_162","b_164"), gs_bands)
nir     <- intersect(c("b_232","b_234","b_242","b_253","b_258","b_259","b_264",
                       "b_267","b_268","b_280","b_281","b_285"), gs_bands)

```{r Calculating Vegetation indices for gs}

ponder_gs <- ponder_gs %>%
  mutate(

    # NDVI variants (canopy vigor / greenness proxies)
    ndvi_gs1 = (b_232 - b_118) / (b_232 + b_118),
    ndvi_gs2 = (b_285 - b_127) / (b_285 + b_127),

    # NDRE variants (red-edge sensitivity; often tracks physiology better than NDVI)
    ndre_gs1 = (b_232 - b_162) / (b_232 + b_162),
    ndre_gs2 = (b_285 - b_164) / (b_285 + b_164),

    # CIgreen-like (NIR/Green - 1)
    ci_green_gs1 = (b_232 / b_76) - 1,
    ci_green_gs2 = (b_285 / b_70) - 1,

    # CIred-edge-like (NIR/RedEdge - 1)
    ci_re_gs1 = (b_232 / b_162) - 1,
    ci_re_gs2 = (b_285 / b_164) - 1,

    # Simple Ratios
    sr_gs1 = b_232 / b_118,
    sr_gs2 = b_285 / b_127,

    # NIR / RedEdge ratios
    srre_gs1 = b_232 / b_162,
    srre_gs2 = b_285 / b_164,

    # Blue–Green pigment proxy (stress / pigment balance-ish)
    nd_blue_green_gs1 = (b_76 - b_23) / (b_76 + b_23),
    nd_blue_green_gs2 = (b_70 - b_19) / (b_70 + b_19),

    # Red-edge slope proxy (edge “strength”)
    re_slope_gs1 = (b_164 - b_162) / (b_164 + b_162)

  )

```

```{r Heat gs indices}
index_cols_gs <- c(
  "ndvi_gs1","ndvi_gs2",
  "ndre_gs1","ndre_gs2",
  "ci_green_gs1","ci_green_gs2",
  "ci_re_gs1","ci_re_gs2",
  "sr_gs1","sr_gs2",
  "srre_gs1","srre_gs2",
  "nd_blue_green_gs1","nd_blue_green_gs2",
  "re_slope_gs1"
)

spearman_idx_gs <- ponder_gs %>%
  summarise(across(
    all_of(index_cols_gs),
    ~cor(.x, gs, method = "spearman", use = "complete.obs")
  )) %>%
  pivot_longer(everything(),
               names_to = "index",
               values_to = "rho")

gs_heat <- ggplot(spearman_idx_gs, aes(x = index, y = "gs", fill = rho)) +
  geom_tile() +
  geom_text(aes(label = sprintf("%.2f", rho)), size = 4) +
  scale_fill_viridis_c(option = "plasma", limits = c(-1,1), name = "Spearman ρ") +
  labs(x = "Vegetation indices", y = NULL) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid = element_blank()
  )
```

```{r Calculating Veg indices for Chl a}

ponder_ca <- ponder_ca %>%
  mutate(

  # NDVI variants (chlorophyll / canopy vigor)
  ndvi_ca1 = (b_252 - b_123) / (b_252 + b_123),
  ndvi_ca2 = (b_271 - b_139) / (b_271 + b_139),

  # NDRE variants (chlorophyll content, nitrogen status)
  ndre_ca1 = (b_252 - b_151) / (b_252 + b_151),
  ndre_ca2 = (b_271 - b_155) / (b_271 + b_155),
  ndre_ca3 = (b_280 - b_156) / (b_280 + b_156),

  # Chlorophyll Index Green (approximate version)
  ci_green1 = (b_252 / b_92) - 1,
  ci_green2 = (b_271 / b_92) - 1,

  # Chlorophyll Index Red Edge
  ci_re1 = (b_252 / b_155) - 1,
  ci_re2 = (b_271 / b_155) - 1,

  # Simple Ratio indices (often correlate with chlorophyll)
  sr_ca1 = b_252 / b_123,
  sr_ca2 = b_271 / b_139,

  # Blue-Red chlorophyll absorption ratio
  br_chl1 = b_34 / b_123,
  br_chl2 = b_48 / b_139,

  # Red-Edge position proxy
  rep_proxy = (b_174 - b_151) / (b_174 + b_151)

  )

```

```{r Heat ca indices}
index_cols <- c(
"ndvi_ca1","ndvi_ca2",
"ndre_ca1","ndre_ca2","ndre_ca3",
"ci_green1","ci_green2",
"ci_re1","ci_re2",
"sr_ca1","sr_ca2",
"br_chl1","br_chl2",
"rep_proxy"
)

spearman_idx <- ponder_ca %>%
  summarise(across(
    all_of(index_cols),
    ~cor(.x, ca, method = "spearman", use = "complete.obs")
  )) %>%
  pivot_longer(everything(),
               names_to = "index",
               values_to = "rho")

ca_heat <- ggplot(spearman_idx, aes(x = index, y = "gs", fill = rho)) +
  geom_tile() +
  geom_text(aes(label = sprintf("%.2f", rho)), size = 4) +
  scale_fill_viridis_c(option = "plasma", limits = c(-1,1), name = "Spearman ρ") +
  labs(x = "Vegetation indices", y = NULL) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid = element_blank())

ggsave(filename = "../Output/heat_ca_spearm.png",
       plot = ca_heat,
       )

```

```{r Calculating Veg indices for pn}
ponder_pn <- ponder_pn %>%
  mutate(

    # --- NDVI family (NIR vs Red) ---
    ndvi_pn1 = (b_253 - b_136) / (b_253 + b_136),
    ndvi_pn2 = (b_281 - b_138) / (b_281 + b_138),
    ndvi_pn3 = (b_242 - b_136) / (b_242 + b_136),

    # --- NDRE family (NIR vs RedEdge) ---
    ndre_pn1 = (b_253 - b_164) / (b_253 + b_164),
    ndre_pn2 = (b_281 - b_162) / (b_281 + b_162),
    ndre_pn3 = (b_242 - b_172) / (b_242 + b_172),

    # --- CIgreen-like (NIR / Green - 1) ---
    ci_green_pn1 = (b_253 / b_74) - 1,   # 906 / 546
    ci_green_pn2 = (b_242 / b_70) - 1,   # 884 / 538

    # --- CIred-edge-like (NIR / RedEdge - 1) ---
    ci_re_pn1 = (b_253 / b_164) - 1,     # 906 / 727
    ci_re_pn2 = (b_281 / b_162) - 1,     # 962 / 723

    # --- Simple Ratio families (often correlate with productivity) ---
    sr_nir_red_pn1 = b_253 / b_136,
    sr_nir_re_pn1  = b_253 / b_164,

    # --- Blue/Green pigment proxies (photosynthesis is pigment-driven) ---
    nd_blue_green1 = (b_74 - b_22) / (b_74 + b_22),   # 546 vs 442
    nd_blue_green2 = (b_70 - b_16) / (b_70 + b_16),   # 538 vs 430

    # --- Red-edge slope / “edge strength” proxies ---
    re_slope_pn1 = (b_174 - b_149) / (b_174 + b_149), # 747 vs 697
    re_slope_pn2 = (b_173 - b_162) / (b_173 + b_162)  # 745 vs 723
  )
```

```{r Heat spear pn}
index_cols_pn <- c(
  "ndvi_pn1",
  "ndvi_pn2",
  "ndvi_pn3",
  "ndre_pn1",
  "ndre_pn2",
  "ndre_pn3",
  "ci_green_pn1",
  "ci_green_pn2",
  "ci_re_pn1",
  "ci_re_pn2",
  "sr_nir_red_pn1",
  "sr_nir_re_pn1",
  "nd_blue_green1",
  "nd_blue_green2",
  "re_slope_pn1",
  "re_slope_pn2"
)

spearman_idx <- ponder_pn %>%
  summarise(across(
    all_of(index_cols_pn),
    ~cor(.x, pn, method = "spearman", use = "complete.obs")
  )) %>%
  pivot_longer(everything(),
               names_to = "index",
               values_to = "rho")

pn_heat <- ggplot(spearman_idx, aes(x = index, y = "gs", fill = rho)) +
  geom_tile() +
  geom_text(aes(label = sprintf("%.2f", rho)), size = 4) +
  scale_fill_viridis_c(option = "plasma", limits = c(-1,1), name = "Spearman ρ") +
  labs(x = "Vegetation indices", y = NULL) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid = element_blank())
```

```{r Calculating Veg indices for ETR}
ponder_etr <- ponder_etr %>%
  mutate(

  # NDVI family (NIR vs Red)
  ndvi_etr1 = (b_253 - b_113)/(b_253 + b_113),
  ndvi_etr2 = (b_281 - b_113)/(b_281 + b_113),
  ndvi_etr3 = (b_238 - b_113)/(b_238 + b_113),

  # NDRE family (NIR vs RedEdge)
  ndre_etr1 = (b_253 - b_164)/(b_253 + b_164),
  ndre_etr2 = (b_238 - b_162)/(b_238 + b_162),
  ndre_etr3 = (b_281 - b_173)/(b_281 + b_173),

  # CIgreen-like
  ci_green_etr1 = (b_253/b_74) - 1,
  ci_green_etr2 = (b_238/b_74) - 1,

  # CIred-edge-like
  ci_re_etr1 = (b_253/b_164) - 1,
  ci_re_etr2 = (b_281/b_162) - 1,

  # Simple ratios
  sr_nir_red_etr1 = b_253/b_113,
  sr_nir_re_etr1  = b_253/b_164,

  # Blue-green pigment proxies
  nd_blue_green_etr1 = (b_74 - b_6)/(b_74 + b_6),
  nd_blue_green_etr2 = (b_74 - b_42)/(b_74 + b_42),

  # Red-edge slope proxies
  re_slope_etr1 = (b_174 - b_145)/(b_174 + b_145),
  re_slope_etr2 = (b_173 - b_162)/(b_173 + b_162)

  )
```

```{r Heat spear etr}
index_cols_etr <- c(
  "ndvi_etr1",
  "ndvi_etr2",
  "ndvi_etr3",
  "ndre_etr1",
  "ndre_etr2",
  "ndre_etr3",
  "ci_green_etr1",
  "ci_green_etr2",
  "ci_re_etr1",
  "ci_re_etr2",
  "sr_nir_red_etr1",
  "sr_nir_re_etr1",
  "nd_blue_green_etr1",
  "nd_blue_green_etr2",
  "re_slope_etr1",
  "re_slope_etr2"
)

spearman_idx <- ponder_etr %>%
  summarise(across(
    all_of(index_cols_etr),
    ~cor(.x, etr, method = "spearman", use = "complete.obs")
  )) %>%
  pivot_longer(everything(),
               names_to = "index",
               values_to = "rho")

etr_heat <- ggplot(spearman_idx, aes(x = index, y = "gs", fill = rho)) +
  geom_tile() +
  geom_text(aes(label = sprintf("%.2f", rho)), size = 4) +
  scale_fill_viridis_c(option = "plasma", limits = c(-1,1), name = "Spearman ρ") +
  labs(x = "Vegetation indices", y = NULL) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid = element_blank())

```
```{r Calculating Veg indices for Chl B}
ponder_cb <- ponder_cb %>%
  mutate(

    # NDVI family (NIR vs Red)
    ndvi_cb1 = (b_252 - b_123) / (b_252 + b_123),
    ndvi_cb2 = (b_271 - b_123) / (b_271 + b_123),

    # NDRE family (NIR vs RedEdge)
    ndre_cb1 = (b_252 - b_155) / (b_252 + b_155),
    ndre_cb2 = (b_271 - b_157) / (b_271 + b_157),
    ndre_cb3 = (b_280 - b_161) / (b_280 + b_161),

    # CIgreen-like (NIR / Green - 1)  [you only have one green band ~540]
    ci_green_cb1 = (b_252 / b_71) - 1,
    ci_green_cb2 = (b_271 / b_71) - 1,

    # CIred-edge-like (NIR / RedEdge - 1)
    ci_re_cb1 = (b_252 / b_155) - 1,
    ci_re_cb2 = (b_271 / b_155) - 1,

    # Simple ratios (often correlate with chlorophyll)
    sr_nir_red_cb1 = b_252 / b_123,
    sr_nir_re_cb1  = b_252 / b_155,

    # Blue/Red absorption ratios (pigment proxies)
    br_cb1 = b_34 / b_123,
    br_cb2 = b_54 / b_123,

    # Blue/Green pigment proxy
    nd_blue_green_cb1 = (b_71 - b_18) / (b_71 + b_18),
    nd_blue_green_cb2 = (b_71 - b_54) / (b_71 + b_54),

    # Red-edge slope / strength proxies (using available RE bands)
    re_slope_cb1 = (b_161 - b_155) / (b_161 + b_155),
    re_slope_cb2 = (b_159 - b_155) / (b_159 + b_155)

  )
```

```{r Heat spear Chl B}
index_cols_cb <- c(
  "ndvi_cb1",
  "ndvi_cb2",
  "ndre_cb1",
  "ndre_cb2",
  "ndre_cb3",
  "ci_green_cb1",
  "ci_green_cb2",
  "ci_re_cb1",
  "ci_re_cb2",
  "sr_nir_red_cb1",
  "sr_nir_re_cb1",
  "br_cb1",
  "br_cb2",
  "nd_blue_green_cb1",
  "nd_blue_green_cb2",
  "re_slope_cb1",
  "re_slope_cb2"
)

spearman_idx_cb <- ponder_cb %>%
  summarise(across(
    all_of(index_cols_cb),
    ~cor(.x, cb, method = "spearman", use = "complete.obs")
  )) %>%
  pivot_longer(everything(),
               names_to = "index",
               values_to = "rho")

cb_heat <- ggplot(spearman_idx_cb, aes(x = index, y = "gs", fill = rho)) +
  geom_tile() +
  geom_text(aes(label = sprintf("%.2f", rho)), size = 4) +
  scale_fill_viridis_c(option = "plasma", limits = c(-1,1), name = "Spearman ρ") +
  labs(x = "Vegetation indices", y = NULL) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid = element_blank())

```
```{r Veg Indices for Cx_C}
ponder_cx_c <- ponder_cx_c %>%
  mutate(

    # --- NDVI / chlorophyll–structure baseline ---
    ndvi_cx1 = (b_271 - b_123) / (b_271 + b_123),
    ndvi_cx2 = (b_281 - b_139) / (b_281 + b_139),

    # --- NDRE (pigment / red-edge sensitivity) ---
    ndre_cx1 = (b_271 - b_155) / (b_271 + b_155),
    ndre_cx2 = (b_281 - b_156) / (b_281 + b_156),
    ndre_cx3 = (b_253 - b_151) / (b_253 + b_151),

    # --- CIgreen-like (NIR / Green - 1) ---
    # You have a green band at ~583 (b_92)
    ci_green_cx1 = (b_271 / b_92) - 1,
    ci_green_cx2 = (b_281 / b_92) - 1,

    # --- CIred-edge-like (NIR / RedEdge - 1) ---
    ci_re_cx1 = (b_271 / b_155) - 1,
    ci_re_cx2 = (b_281 / b_155) - 1,

    # --- Blue/Green pigment proxies (carotenoids live in the blue) ---
    nd_blue_green_cx1 = (b_92 - b_18) / (b_92 + b_18),
    nd_blue_green_cx2 = (b_92 - b_42) / (b_92 + b_42),

    # --- Blue/Red pigment balance proxies ---
    br_cx1 = b_36 / b_123,   # blue-ish / red
    br_cx2 = b_49 / b_139,   # blue-ish / deep red

    # --- Red-edge slope proxies (edge strength / pigment transition) ---
    re_slope_cx1 = (b_174 - b_146) / (b_174 + b_146),
    re_slope_cx2 = (b_173 - b_151) / (b_173 + b_151)

  )
```

```{r Heatmap Cx_C}
index_cols_cx_c <- c(
  "ndvi_cx1",
  "ndvi_cx2",
  "ndre_cx1",
  "ndre_cx2",
  "ndre_cx3",
  "ci_green_cx1",
  "ci_green_cx2",
  "ci_re_cx1",
  "ci_re_cx2",
  "nd_blue_green_cx1",
  "nd_blue_green_cx2",
  "br_cx1",
  "br_cx2",
  "re_slope_cx1",
  "re_slope_cx2"
)

spearman_idx_cxc <- ponder_cx_c %>%
  summarise(across(
    all_of(index_cols_cx_c),
    ~cor(.x, cx_c, method = "spearman", use = "complete.obs")
  )) %>%
  pivot_longer(everything(),
               names_to = "index",
               values_to = "rho")

cb_heat <- ggplot(spearman_idx_cxc, aes(x = index, y = "gs", fill = rho)) +
  geom_tile() +
  geom_text(aes(label = sprintf("%.2f", rho)), size = 4) +
  scale_fill_viridis_c(option = "plasma", limits = c(-1,1), name = "Spearman ρ") +
  labs(x = "Vegetation indices", y = NULL) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid = element_blank())

```

```{r}
ggplot(data = selected_bands_wl,
       aes(x = as.factor(wavelength),
           y = coefficient
           )) +
  geom_col(width = 2) +
  geom_hline(yintercept = 0) +
  geom_vline(xintercept = 450, color= "blue") +
  geom_vline(xintercept = 550, color = "green") +
  geom_vline(xintercept = 660, color = "red") +
  geom_line(aes(group = 1)) +
  facet_wrap(~ variable, scales = "free_y") +
  
  labs(title = element_text("Importance Distribution Across Spectrum"),
       x = element_text("Wavelength (nm)"),
       y = element_text("Coefficient Value"))
  

  
```

#5) Regressions

```{r Setting up Dataframe to Use Only Desired Bands Selected by Elastic Net.}
long_ponder <- clean_ponder %>%
  pivot_longer(cols = contains("b_"),
               values_to = "reflect",
               names_to = "band")%>%
  pivot_longer(cols = c("net_photo":"cx_c"), 
               names_to = "variable")

write_csv(long_ponder, file = "../Output/bands.csv")

regression <- long_ponder %>% 
  inner_join(selec, 
            by = c("band", "variable"))

write_csv(regression, file = "../Output/bands_filt.csv")

```

```{r Subset of Variables 74 DAP}

ca74 <- regression %>%
  filter(variable == "ca" & dap == 74) %>%
  select(band:value, reflect) %>% 
  pivot_wider(
    names_from = band,
    values_from = reflect,
    values_fn = mean   # or median
  ) %>%
  rename(ca74 = value) %>%
  select(-variable)

cx_c74 <- regression %>%
  filter(variable == "cx_c" & dap == 74) %>%
  select(band:value, reflect) %>% 
  pivot_wider(
    names_from = band,
    values_from = reflect,
    values_fn = mean   # or median
  ) %>%
  rename(cx_c74 = value) %>%
  select(-variable)

cb74 <- regression %>%
  filter(variable == "cb" & dap == 74) %>%
  select(band:value, reflect) %>% 
  pivot_wider(
    names_from = band,
    values_from = reflect,
    values_fn = mean   # or median
  ) %>%
  rename(cb74 = value) %>%
  select(-variable)

etr74 <- regression %>%
  filter(variable == "etr" & dap == 74) %>%
  select(band:value, reflect) %>% 
  pivot_wider(
    names_from = band,
    values_from = reflect,
    values_fn = mean   # or median
  ) %>%
  rename(etr74 = value) %>%
  select(-variable)


gs74 <- regression %>%
  filter(variable == "gs" & dap == 74) %>%
  select(band:value, reflect) %>% 
  pivot_wider(
    names_from = band,
    values_from = reflect,
    values_fn = mean   # or median
  ) %>%
  rename(gs74 = value) %>%
  select(-variable)


net_photo74 <- regression %>%
  filter(variable == "net_photo" & dap == 74) %>%
  select(band:value, reflect) %>% 
  pivot_wider(
    names_from = band,
    values_from = reflect,
    values_fn = mean   # or median
  ) %>%
  rename(net_photo74 = value) %>%
  select(-variable)

```

```{r Subset of Variables 81 DAP}


ca81 <- regression %>%
  filter(variable == "ca" & dap == 81) %>%
  select(band:value, reflect) %>% 
  pivot_wider(
    names_from = band,
    values_from = reflect,
    values_fn = mean   # or median
  ) %>%
  rename(ca81 = value) %>%
  select(-variable)

cx_c81 <- regression %>%
  filter(variable == "cx_c" & dap == 81) %>%
  select(band:value, reflect) %>% 
  pivot_wider(
    names_from = band,
    values_from = reflect,
    values_fn = mean   # or median
  ) %>%
  rename(cx_c81 = value) %>%
  select(-variable)

cb81 <- regression %>%
  filter(variable == "cb" & dap == 81) %>%
  select(band:value, reflect) %>% 
  pivot_wider(
    names_from = band,
    values_from = reflect,
    values_fn = mean   # or median
  ) %>%
  rename(cb81 = value) %>%
  select(-variable)

etr81 <- regression %>%
  filter(variable == "etr" & dap == 81) %>%
  select(band:value, reflect) %>% 
  pivot_wider(
    names_from = band,
    values_from = reflect,
    values_fn = mean   # or median
  ) %>%
  rename(etr81 = value) %>%
  select(-variable)


gs81 <- regression %>%
  filter(variable == "gs" & dap == 81) %>%
  select(band:value, reflect) %>% 
  pivot_wider(
    names_from = band,
    values_from = reflect,
    values_fn = mean   # or median
  ) %>%
  rename(gs81 = value) %>%
  select(-variable)


net_photo81 <- regression %>%
  filter(variable == "net_photo" & dap == 81) %>%
  select(band:value, reflect) %>% 
  pivot_wider(
    names_from = band,
    values_from = reflect,
    values_fn = mean   # or median
  ) %>%
  rename(net_photo81 = value) %>%
  select(-variable)

```

```{r Subset of Variables 109 DAP}


ca109 <- regression %>%
  filter(variable == "ca" & dap == 109) %>%
  select(band:value, reflect) %>% 
  pivot_wider(
    names_from = band,
    values_from = reflect,
    values_fn = mean   # or median
  ) %>%
  rename(ca109 = value) %>%
  select(-variable)

cx_c109 <- regression %>%
  filter(variable == "cx_c" & dap == 109) %>%
  select(band:value, reflect) %>% 
  pivot_wider(
    names_from = band,
    values_from = reflect,
    values_fn = mean   # or median
  ) %>%
  rename(cx_c109 = value) %>%
  select(-variable)

cb109 <- regression %>%
  filter(variable == "cb" & dap == 109) %>%
  select(band:value, reflect) %>% 
  pivot_wider(
    names_from = band,
    values_from = reflect,
    values_fn = mean   # or median
  ) %>%
  rename(cb109 = value) %>%
  select(-variable)

etr109 <- regression %>%
  filter(variable == "etr" & dap == 109) %>%
  select(band:value, reflect) %>% 
  pivot_wider(
    names_from = band,
    values_from = reflect,
    values_fn = mean   # or median
  ) %>%
  rename(etr109 = value) %>%
  select(-variable)


gs109 <- regression %>%
  filter(variable == "gs" & dap == 109) %>%
  select(band:value, reflect) %>% 
  pivot_wider(
    names_from = band,
    values_from = reflect,
    values_fn = mean   # or median
  ) %>%
  rename(gs109 = value) %>%
  select(-variable)


net_photo109 <- regression %>%
  filter(variable == "net_photo" & dap == 109) %>%
  select(band:value, reflect) %>% 
  pivot_wider(
    names_from = band,
    values_from = reflect,
    values_fn = mean   # or median
  ) %>%
  rename(net_photo109 = value) %>%
  select(-variable)

```

```{r Setting up Models, Hyperparameters, and Evaluation Metrics}

# Evaluation function

evaluate_model <- function(model, dataset, target_name, dataset_name) {
  predicoes <- predict(model, dataset)
  obs <- dataset[[target_name]]
  data.frame(
    Dataset = dataset_name,
    MAE = mean(abs(predicoes - obs)),
    MSE = mean((predicoes - obs)^2),
    RMSE = sqrt(mean((predicoes - obs)^2)),
    R2 = cor(predicoes, obs)^2
  )
}

# Dataset names

datasets <- list(
  ca74 = ca74, ca81 = ca81, ca109 = ca109,
  cb74 = cb74, cb81 = cb81, cb109 = cb109,
  cx_c74 = cx_c74, cx_c81 = cx_c81, cx_c109 = cx_c109,
  etr74 = etr74, etr81 = etr81, etr109 = etr109,
  gs74 = gs74, gs81 = gs81, gs109 = gs109,
  net_photo74 = net_photo74, net_photo81 = net_photo81, net_photo109 = net_photo109
)

# Controle de treinamento
ctrl <- trainControl(method = "repeatedcv", number = 5, repeats = 5, search = "grid")

# Grades de hiperparâmetros
tune_grid_rf <- expand.grid(mtry = c(2, 3, 4, 5), splitrule = c("variance", "extratrees"), min.node.size = c(1, 5, 10))
tune_grid_svm <- expand.grid(C = c(0.1, 1, 10, 100), sigma = c(0.001, 0.01, 0.1, 1))
tune_grid_knn <- expand.grid(kmax = c(3, 5, 7, 9, 11), distance = c(1, 2), kernel = c("rectangular", "triangular", "epanechnikov"))
tune_grid_xgbTree <- expand.grid(nrounds = c(500, 1000, 1500), max_depth = c(2, 4, 6), eta = c(0.01, 0.1, 0.3),
                                 gamma = c(0, 0.1, 0.2), colsample_bytree = c(0.4, 0.6, 0.8, 1), min_child_weight = c(1, 3, 5), subsample = 1)
```

```{r Random Forest Regression Model}
resultados_rf <- data.frame()

for (nome_dataset in names(datasets)) {
  
  df <- datasets[[nome_dataset]] %>%
    na.omit() %>%
    mutate(across(starts_with("b_"), ~ suppressWarnings(as.numeric(as.character(.))))) %>%
    select(where(~ !(all(is.na(.)) & is.numeric(.))))   # DROP BAD COLUMNS
  
  target <- names(df)[1]
  
  set.seed(123)
  inTraining <- createDataPartition(df[[target]], p = 0.8, list = FALSE)
  treino <- df[inTraining,]
  teste <- df[-inTraining,]
  
  model_rf <- train(
    formula(paste(target, "~ .")),
    data = treino,
    method = "ranger",
    trControl = ctrl,
    tuneGrid = tune_grid_rf,
    importance = "impurity"
  )
  
  metricas_treino <- evaluate_model(model_rf, treino, target, paste(nome_dataset, "- Treino"))
  metricas_teste  <- evaluate_model(model_rf, teste,  target, paste(nome_dataset, "- Teste"))
  
  resultados_rf <- bind_rows(resultados_rf, metricas_treino, metricas_teste)
}

print(resultados_rf)


```

```{r Support Vector Machine Regression Model}

resultados_svm <- data.frame()
for (nome_dataset in names(datasets)) {
  df <- datasets[[nome_dataset]] %>% na.omit()
  target <- names(df)[1]
  set.seed(123)
  inTraining <- createDataPartition(df[[target]], p = 0.8, list = FALSE)
  treino <- df[inTraining,]
  teste <- df[-inTraining,]
  model_svm <- train(
    formula(paste(target, "~ .")),
    data = treino,
    method = "svmRadial",
    preProcess = c("center", "scale"),
    trControl = ctrl,
    tuneGrid = tune_grid_svm
  )
  metricas_treino <- evaluate_model(model_svm, treino, target, paste(nome_dataset, "- Treino"))
  metricas_teste <- evaluate_model(model_svm, teste, target, paste(nome_dataset, "- Teste"))
  resultados_svm <- bind_rows(resultados_svm, metricas_treino, metricas_teste)
}
print(resultados_svm)
```

```{r K Nearest Neighbor Regression Model}

resultados_knn <- data.frame()
for (nome_dataset in names(datasets)) {
  df <- datasets[[nome_dataset]] %>% na.omit()
  target <- names(df)[1]
  set.seed(123)
  inTraining <- createDataPartition(df[[target]], p = 0.8, list = FALSE)
  treino <- df[inTraining,]
  teste <- df[-inTraining,]
  model_knn <- train(
    formula(paste(target, "~ .")),
    data = treino,
    method = "kknn",
    preProcess = c("center", "scale"),
    trControl = ctrl,
    tuneGrid = tune_grid_knn
  )
  metricas_treino <- evaluate_model(model_knn, treino, target, paste(nome_dataset, "- Treino"))
  metricas_teste <- evaluate_model(model_knn, teste, target, paste(nome_dataset, "- Teste"))
  resultados_knn <- bind_rows(resultados_knn, metricas_treino, metricas_teste)
}
print(resultados_knn)
```


```{r XGBoost Regression Model}

resultados_xgb <- data.frame()
for (nome_dataset in names(datasets)) {
  df <- datasets[[nome_dataset]] %>% na.omit()
  target <- names(df)[1]
  set.seed(123)
  inTraining <- createDataPartition(df[[target]], p = 0.8, list = FALSE)
  treino <- df[inTraining,]
  teste <- df[-inTraining,]
  model_xgb <- train(
    formula(paste(target, "~ .")),
    data = treino,
    method = "xgbTree",
    trControl = ctrl,
    tuneGrid = tune_grid_xgbTree,
    verbose = FALSE
  )
  metricas_treino <- evaluate_model(model_xgb, treino, target, paste(nome_dataset, "- Treino"))
  metricas_teste <- evaluate_model(model_xgb, teste, target, paste(nome_dataset, "- Teste"))
  resultados_xgb <- bind_rows(resultados_xgb, metricas_treino, metricas_teste)
}
print(resultados_xgb)
```

```{r Wrap-up And Exporting Results}
# Adicionar uma coluna indicando o modelo correspondente
resultados_rf$modelo <- "Random Forest"
resultados_svm$modelo <- "SVM Radial"
resultados_knn$modelo <- "kNN"
#resultados_xgb$modelo <- "XGBoost "

# Juntar todas as tabelas em um único dataframe
resultados_finais <- bind_rows(
  resultados_rf,
  resultados_svm,
  resultados_knn,
  #resultados_xgb
)

# Exibir a tabela final consolidada
print(resultados_finais)

# Se quiser salvar em CSV:
write.csv(resultados_finais, "../Output/resultados.csv", row.names = FALSE)
```


